
```{r}
#| output: asis
#| echo: false
model=gsub("~","",mrs@metadata$model)
level=gsub(model,"",index)
cat ("\n\n#### ", level, "\n\n")
```

::: panel-tabset

##### DMPS

::: panel-tabset

###### Differentially methylated probes

```{r}
dmps<-getDMPs(mrs,index)
readr::write_tsv(dmps,file=paste0("dmps_", mrs@metadata$sva, mrs@metadata$model, "_", mrs@metadata$method, "_", level, "_vs_", mrs@metadata$intercept, ".tsv"))
nbdmps<-dmps %>% filter(adj.P.Val<0.05) %>% nrow()
```

Number of significant dmps = `r nbdmps`

```{r}
dmps %>% dplyr::filter(adj.P.Val<0.05) %>% 
  arrange(adj.P.Val) %>% head(100) %>%
  DT::datatable(extensions = 'Buttons', options = list(
                           dom = 'Blfrtip',
                           buttons = c('copy', 'csv')))
```

::: {.callout-tip}
This table provides a comprehensive overview of DMPs, including statistical metrics, genomic coordinates, strand information, and functional annotations
:::

::: {.callout-note collapse="true"}

* deltabetas: Difference in methylation levels between experimental groups (e.g., high vs. low).
* AveExp: Average methylation level across experimental groups.
* t: t-statistic indicating the magnitude and direction of the difference in methylation levels.
* P.value: p-value from statistical testing, indicating the significance of the observed difference.
* adj.P.Value: Adjusted p-value after correction for multiple testing.
* B: Estimate of methylation difference.
* Stdev: Standard deviation of methylation levels.
* Goodness: Measure of model fit or goodness of fit statistic.

:::

###### QQplot

```{r}
qqman::qq(dmps$P.Value,main=paste0("QQ plot: ","lambda=", mrs@lambda[[index]]))
```

::: {.callout-tip}
## QQplot

Quantile-Quantile (QQ) plot is a graphical tool used to assess whether the observed distribution of methylation data conforms to the expected distribution, typically assumed to be normal. In the context of methylation analysis, QQ plots help evaluate the goodness of fit of statistical models and identify potential deviations from expected distributions, such as skewness or outliers.
:::


###### Volcano Plot
  
```{r}
volcano( dmps )
```

::: {.callout-tip}
## Volcano Plot

Points located in the left or right sides of the plot represent DMPs with large deltabetas, indicating substantial differences in methylation levels between groups. Points located to the upper regions of the plot represent DMPs with low adjusted p-values, indicating statistical significance.
:::

###### Circos Plot

```{r}  
circosplot( get_dmps_ranges(mrs,group=index,q=0.05), params$genome )
```

::: {.callout-tip}
## Circos plot

The Circos plot provides a holistic view of the genomic landscape, highlighting spatial patterns and chromosome-wide distributions of methylation changes.
:::

:::

##### DMRS

```{r}
#| output: false
result <- get_results(mrs, index, tools=params$dmrtools, genome=params$genome, mvals=mvals )
```

::: panel-tabset

###### Summary

::: {.callout-note}
Differentially Methylated Regions (DMRs) are contiguous genomic regions where the methylation levels exhibit statistically significant differences between experimental groups. Unlike individual differentially methylated positions (DMPs), DMRs provide a more comprehensive view of methylation changes by considering spatial continuity and local correlations in methylation patterns. DMR analysis is a crucial step in methylation studies for identifying biologically relevant regions associated with the studied conditions or phenotypes.
:::

Number of significant dmrs identified by tools

```{r}
result <- result %>% resultsToLong()
result %>% filter(fdr<0.05, no.cpgs>1) %>% group_by(ID,dmrtool) %>% summarise(n=n()) %>% pull(dmrtool) %>% table()
saveRDS(result,"result.debug.rds")
```


###### Volcano
  
```{r}
saveRDS(result,file="result.rds")

genes_of_interest <- c("Casp7", "Dpysl2", "Gata6", "Ifi27", "Mir143-145",
                        "Pcp4l1", "Pphln1", "Prune1", "Runx1", "Sox5")

my_volcanomean <- function(df, genes_of_interest = NULL){
  if(is.null(df) | nrow(df)<1) { return(NULL) }
  
  volcano_data <- df %>% 
    group_by(dmrtool, ID, fdr, genesUniq) %>% 
    summarise(no.cpgs = n(), mean.deltabetas = mean(deltabetas), .groups = 'drop') %>%
    mutate(
      sig_status = ifelse(fdr > 0.05, "3notsig", ifelse(mean.deltabetas > 0, "1up", "2down")),
      label = ifelse(!is.null(genes_of_interest) & genesUniq %in% genes_of_interest, genesUniq, NA)
    )
  
  volcano <- volcano_data %>%
    ggplot(aes(x = mean.deltabetas, y = -log10(fdr))) +
    geom_point(aes(color = sig_status), alpha = 0.5) +
    scale_colour_manual(values = c("blue", "red", "lightgray")) +
    geom_hline(yintercept = -log10(0.05), linetype = "dashed", col = "red") +
    geom_label_repel(
      aes(label = label),
      na.rm = TRUE,
      size = 3,
      max.overlaps = Inf,
      box.padding = 0.4,
      label.padding = 0.2,
      segment.color = "grey40",
      segment.size = 0.4,
      fontface = "italic",
      force = 1,              # force de répulsion entre labels (augmenter si entassés)
      #point.padding = 0.8,
      nudge_y = 50
    ) +
    guides(color = "none") + 
    facet_grid(. ~ dmrtool) +
    theme_minimal() + 
    ggtitle("A : mean deltabetas")
  
  return(volcano)
}

my_volcanomean(result,genes_of_interest)
```

::: {.callout-note}
Volcano plot, points located in the left or right sides of the plot represent DMRs with large mean deltabetas, indicating substantial differences in methylation levels between groups. Points located to the upper regions of the plot represent DMPs with low adjusted p-values, indicating statistical significance.
:::

###### DMR table  
  
```{r}
result <- left_join(dmps %>% filter(adj.P.Val<0.05), result, by=colnames(dmps))
readr::write_tsv(result,file=paste0("dmrs_", mrs@metadata$sva, mrs@metadata$model, "_", mrs@metadata$method, "_", level, "_vs_", mrs@metadata$intercept, ".tsv"))
result %>% dplyr::filter(fdr<0.05, no.cpgs>1) %>% 
  arrange(fdr) %>% head(100) %>%
  DT::datatable(extensions = 'Buttons', options = list(
                           dom = 'Blfrtip',
                           buttons = c('copy', 'csv')))
```

::: {.callout-note}
The DMR table extends the information provided in the DMPs table by annotating regions of contiguous methylation changes.
:::

###### Tools comparison

::: panel-tabset

####### Venn

```{r}
my_venn(result %>% filter(fdr<0.05, no.cpgs>1))
```

::: {.callout-note}
Regions in the overlapping areas of the Venn diagram represent Differentilly Methylated Regions that are commonly identified by multiple tools (contain at least one common CpG), suggesting robustness and consistency in their identification.
:::

####### Density
  
```{r}
my_density(result)
```

::: {.callout-note}
Density plot of DMR lengths, distribution and variability of region lengths
:::
  
####### Violin

```{r}
my_violin(result)
```

::: {.callout-note}
Similar plot showing DMR length distribution from an other angle.
:::
  
:::

:::

:::