---
title: "Data Loading"
subtitle: "Loading methylation data and creating analysis objects"
params:
  idat_path: "/data/EPI-AGING/files/ILL_20240614_epiaging_mm/idats/"
  sample_sheet_path: "../sampleSheet.csv"
  debug: true
---

This chapter covers the initial steps of loading methylation data from IDAT
files and creating the core analysis objects used throughout the methylkey 
workflow.

```{r setup}
#| echo: false
devtools::load_all("/home/cahaisv/git/methylkey2/")
```

```{r debug}
#| echo: false
#| eval: !expr params$debug
#| results: asis
full_sample_path <- normalizePath(params$sample_sheet_path, mustWork = FALSE)
full_idat_path   <- normalizePath(params$idat_path, mustWork = FALSE)

# On définit une petite fonction pour simplifier
red_bold <- function(x) paste0("<span style='color: red; font-weight: bold;'>", x, "</span>")

cat("\n::: callout-warning\n\n")
cat("\n## Debug Information\n")
cat(red_bold("Working Directory :\n"), getwd(), "\n\n")
cat(red_bold("Sample sheet      :\n"), full_sample_path, "\n\n")
cat(red_bold("IDAT path         :\n"), full_idat_path, "\n\n")
cat("\n:::\n\n---\n\n")
```

## Sample Sheet Preparation

```{r sampleSheet}
# Read sample sheet
sample_sheet <- readr::read_tsv(params$sample_sheet_path) |>
  formatSampleSheet()
DT::datatable(sample_sheet)
```

## Loading IDAT Files

```{r re_idat}
#| eval: !expr "file.exists('meth.rds')"
#| echo: false
meth<-readRDS("meth.rds")
```

```{r idat}
#| eval: !expr "!file.exists('meth.rds') | !file.exists('sdfs.rds')"

# Clock_models
zhoulab_url = "https://github.com/zhou-lab/InfiniumAnnotationV1/raw/main/Anno/"
Clock_models=list(
  Zhou347=file.path(zhoulab_url,"MM285/Clock_Zhou347.rds"))

# Load data with sesame pipeline
meth <- sesame2Betas(
  idat = params$idat_path,
  sampleSheet=sample_sheet,
  prep = "TQCDPB",
  na=0.2, 
  ncore=1,
  Clock_models=Clock_models)

# Save the processed object
saveRDS(meth, "meth.rds")
```

Plateform : `r getPlateform(meth)`

Number of samples : `r nbsamples(meth)`

Number of probes : `r nbprobes(meth)`

::: callout-note
## Generated object

sdfs: This file is generated by sesame it contains the raw data intensities.

meth: This is a summarizedExperiment object that contains betas, sampleSheet 
manifest and metadatas about the analysis.
:::


